\documentclass[aspectratio=1610,8pt]{beamer}

% Standard packages

\usepackage[english]{babel}
%\usepackage[latin1]{inputenc}
%\usepackage{times}
%\usepackage[T1]{fontenc}
\usepackage{fontspec}
\usepackage[]{unicode-math}
\setmathfont{Inconsolata}
\setsansfont{Roboto}

% Setup TikZ

\usepackage{tikz}
\usetikzlibrary{arrows}
\tikzstyle{block}=[draw opacity=0.7,line width=1.4cm]


% Author, Title, etc.

\title{Summation of Series\\Problems 11-20}

\author[Shiv Shankar Dayal]{Shiv Shankar Dayal}

% The main document

\begin{document}
\begin{frame}
  \titlepage
\end{frame}
\begin{frame}{Problem 11}
  \textbf{11.} Find the sum of $n$ terms of the series $1 + 5 + 11 + 19 + \ldots$
\end{frame}
\begin{frame}{Solution of Problem 11}
  \textbf{Solution:} Let
  $$S_n = 1 + 5 + 11 + 19 + \ldots + t_n$$
  $$S_n = 1 + 5 + 11 + 19 + \ldots + t_{n -1} + t_n$$
  Subtracting first term of lower series with second term of upper series, second term of lower series from third term of upper
  series and so on, we get
  $$0 = 1 + [4 + 6 + 8 + \ldots \text{to~}n - 1~\text{terms}] - t_n$$
  $$t_n = 1 + \frac{n - 1}{2}[2.4 + (n - 2)2] = n^2 + n - 1$$
  $$S_n = \sum t_n = \sum n^2 + \sum n - \sum 1$$
  $$= \left[\frac{n(n + 1)(2n + 1)}{6} + \frac{n(n + 1)}{2} - n\right]$$
  $$= \frac{n(n^2 + 3n - 1)}{3}$$
\end{frame}
\begin{frame}{Problem 12}
  \textbf{12.} A sum is distributed among certain number of persons. Second person gets one rupee more than the first, third person
  gets two rupees more than the second, fourth person gets three rupees more than the third and so on. If the first person gets one
  rupee and the last person get $67$ rupees, find the number of persons.
\end{frame}
\begin{frame}{Solution of Problem 12}
  \textbf{Solution:} Given, first person gets $1$ rupees, second person gets $1 + 1 = 2$ rupees, third person gets $2 + 2 = 4$
  rupees, fourth person gets $4 + 3 = 7$ rupees and so on.

  Let there be $n$ persons and $t_n$ be the amount last person gets which is given as $67$ rupees.
  $$S_n = 1 + 2 + 4 + 7 + \ldots + t_n$$
  $$S_n = 1 + 2 + 4 + 7 + \ldots + t_n$$
  Subtracting first term of lower series with second term of upper series, second term of lower series from third term of upper
  series and so on, we get
  $$0 = 1 + [1 + 2 + 3 + \ldots \text{to~}n - 1~\text{terms}] - t_n$$
  $$t_n = 1 + \frac{n(n - 1)}{2} = \frac{n^2 - n + 2}{2} = 67$$
  $$n^2 - n + 132 = 0 \Rightarrow (n - 12)(n + 11) = 0\Rightarrow n = 12, -11$$
  But no. of persons cannot be negative, therefore no. of persons is $12.$
\end{frame}
\begin{frame}{Problem 13}
  \textbf{13.} Natural numbers have been grouped in the following way $1, (2, 3), (4, 5, 6), (7, 8, 9, 10), \ldots$ Show that the
  sum of the numbers in the $n$th group is $\frac{n(n^2 + 1)}{2}.$
\end{frame}
\begin{frame}{Solution of Problem 13}
  \textbf{Solution:} Since $1$st group contains $1$ number, $2$nd group contains $2$ numbers, $3$rd group contains $3$ numbers, and
  so on, therefore, $n$th group will contain $n$ numbers.

  Here, we observe that numbers in each group are in A.P. whose c.d. is $1. (n - 1)$th group will have $(n - 1)$ numbers. Total
  numbers of numbers till the end of $(n - 1)$th group is

  $$N=1 + 2 + 3 +\ldots \text{~for~} (n - 1) \text{~terms}$$
  $$= \frac{n(n - 1)}{2}$$

  Thus, first term of nth group will be $N + 1$ i.e $\frac{n^2 - n + 2}{2}$

  Thus, sum of numbers in $n$th group is

  $$S = \frac{n}{2}\left(2.\frac{n^2 - n + 2}{2} + (n - 1).1\right) = \frac{n(n^2 + 1)}{2}$$
\end{frame}
\begin{frame}{Problem 14}
  \textbf{14.} Find $1 + 3 + 7 + 15 + \ldots$ to $n$ terms.
\end{frame}
\begin{frame}{Solution of Problem 14}
  \textbf{Solution:} $$S = 1 + 3 + 7 + 15 + \ldots + t_n$$
  $$S = 1 + 3 + 7 + 15 + \ldots + t_{n - 1} + t_n$$
  Subtracting first term of lower series with second term of upper series, second term of lower series from third term of upper
  series and so on, we get
  $$0 = [1 + 2 + 4 + 8 + \ldots\text{~to~}n\text{~terms}] - t_n$$
  $$t_n = 2^n - 1$$
  $$S = \sum t_n = [2^1 + 2^2 + \ldots +2^n] - \sum 1 = 2(2^n - 1) - n = 2^{n + 1} - 2 - n$$
\end{frame}
\begin{frame}{Problem 15}
  \textbf{15.} Find $1 + 2x + 3x^2 + 4x^3 + \ldots$ to $n$ terms.
\end{frame}
\begin{frame}{Solution of Problem 15}
  \textbf{Solution:}
  $$S_n = 1 + 2x + 3x^2 + \ldots + n.x^{n- 1}$$
  $$xS_n = x + 2x^2 + 3x^^3 + \ldots + (n - 1)x^{n-1}  + n.x^n$$
  Subtracting first term of lower series with second term of upper series, second term of lower series from third term of upper
  $$(1 - x)S_n = [1 + x + x^2 + \ldots \text{~to~}n\text{~terms}] -nx^n$$
  $$S_n = \frac{1 - x^n}{(1 - x)^2} - \frac{nx^n}{1 - x}$$
\end{frame}
\begin{frame}{Problem 16}
  \textbf{16.} Find $1 + 2.2 + 3.2^2 + 4.3^3 + \ldots + 100.2^{99}$
\end{frame}
\begin{frame}{Solution of Problem 16}
  \textbf{Solution:} $$S = 1 + 2.2 + 3.2^2 + 4.3^3 + \ldots + 100.2^{99}$$
  $$2S = 1.2 + 2.2^2 + 3.2^3 + \ldots + 99.2^{99} + 100.2^{100}$$
  Subtracting first term of lower series with second term of upper series, second term of lower series from third term of upper
  $$-S = 1 + 2 + 2^2 + 2^3 + \ldots + 2^{99} - 100.2^{100}$$
  $$S = 100.2^{100} - (2^{100} - 1)$$
  $$= 99.2^{100} + 1$$
\end{frame}
\begin{frame}{Problem 17}
  \textbf{17.} find $1 + 2^2x + 3^2x^2 + 4^2x^4 + \ldots \text{~to~}\infty, |x| < 1$
\end{frame}
\begin{frame}{Solution of Problem 17}
  \textbf{Solution:}$$S = 1 + 2^2x + 3^2x^2 + 4^2x^3 + \ldots \text{~to~}\infty$$
  $$xS = x + 2^2x^2 + 3^2x^3 + \ldots \text{~to~}\infty$$
  $$(1 - x)S = 1 + 3x + 5x^2 + 7x^3 + \ldots \text{~to~}\infty$$
  Repeating the process again
  $$(1 - x)^2S = 1 + 2x(1 + x + x^2 + \ldots \text{~to~}\infty)$$
  $$S = \frac{1 + \frac{2x}{1 - x}}{(1 - x)^2} = \frac{1 + x}{(1 - x)^3}$$
\end{frame}
\begin{frame}{Problem 18}
  \textbf{18.} If the sum of $n$ terms of a sequence be $2n^2 + 4,$ find its $n$th term. Is this sequence in A.P.?
\end{frame}
\begin{frame}{Solution of Problem 18}
  \textbf{Solution:}$$S_n = 2n^2 + 4 \Rightarrow S_{n - 1} = 2(n - 1)^2 + 4 = 2n^2 -4n +6$$
  $$t_n = S_n - S_{n - 1} = 4n - 6$$
  $$\Rightarrow t_{n - 1} = 4n - 2$$
  $$d = t_n - t_{n - 1} = 4$$
  Since common difference is a constant the sequence is in A.P.
\end{frame}
\begin{frame}{Problem 19}
  \textbf{19.} Find the sum of $n$ terms of the series whose $n$th term is $n(n - 1)(n + 1)$
\end{frame}
\begin{frame}{Solution of Problem 19}
  \textbf{Solution:}$$t_n = n(n - 1)(n + 1) = n^3 - n$$
  $$S_n = \sum n^3 - \sum n = \left(\frac{n(n + 1)}{2}\right)^2 - \frac{n(n + 1)}{2}$$
  $$= \frac{n(n + 1)}{2}\left[\frac{n(n + 1)}{2} - 1\right]$$
  $$= \frac{n(n + 1)(n^2 + n - 2)}{4}$$
\end{frame}
\begin{frame}{Problem 20}
  \textbf{20.} Find the sum of $80$ terms of the series whose $n$th term is $n(n^2 - 1)$
\end{frame}
\begin{frame}{Solution of Problem 20}
  \textbf{Solution:} The $n$th term is same as problem 19. Thus,
  $$S_{80} = \frac{80.81.(80^2 + 80 - 2)}{4} = 10494360$$
\end{frame}
\end{document}
