\documentclass[aspectratio=1610,8pt]{beamer}

% Standard packages

\usepackage[english]{babel}
%\usepackage[latin1]{inputenc}
%\usepackage{times}
%\usepackage[T1]{fontenc}


% Setup TikZ

\usepackage{tikz}
\usetikzlibrary{arrows}
\tikzstyle{block}=[draw opacity=0.7,line width=1.4cm]


% Author, Title, etc.

\title{Problems 31 to 40}

\author[Shiv Shankar Dayal]{Shiv Shankar Dayal}

% The main document

\begin{document}
\begin{frame}
       \titlepage
\end{frame}
\begin{frame}{Choosing numbers}
  \begin{itemize}
    \item When you have to choose three numbers in arithmetic progression it is
      better to choose them as $a - d, a, a + d$
    \item When you have to choose four numbers in arithmetic progression it is
      better to choose them as $a-3d, a - d, a + d, a + 3d$
    \item When you have to choose five numbers in arithmetic progression it is
      better to choose them as $a-2d, a - d, a, a + d, a + 2d$
  \end{itemize}
\end{frame}
\begin{frame}{Problem 31}
  \textbf{31.} The sum of three numbers in A.P. is $27$ and the sum of their
  squares is $293.$ Find the numbers.
\end{frame}
\begin{frame}{Solution of problem 31}
  \textbf{Solution:} Let the three numbers in A.P. be $a - d, a, a + d$\\
  Given, $a - d + a + a + d = 27 \Rightarrow 3a = 27 \therefore a = 9$\\
  and $(a - d)^2 + a^2 + (a + d)^2 = 293$\\
  $\Rightarrow a^2 + d^2 - 2ad + a^2 + a^2 + d^2 + 2ad = 293$\\
  $3a^2 + 2d^2 = 293 \Rightarrow 3.9^2 + 2d^2 = 293$\\
  $243 + 2d^2 = 293 \Rightarrow 2d^2 = 50 \Rightarrow d^2 = 25$\\
  $d = \pm 5$\\
  If $d = -5,$ the three numbers are $14, 9, 4$\\
  If $d = 5,$ the three numbers are $4, 9, 14$
\end{frame}
\begin{frame}{Problem 32}
  \textbf{32.} The sum of four integers in A.P. is $24$ and their product is
  $945.$ Find the numbers.
\end{frame}
\begin{frame}{Solution of problem 32}
  \textbf{Solution:} Let the four numbers be $a - 3d, a - d, a + d, a + 3d$\\
  Given, $a - 3d + a - d + a + d + a + 3d = 24$\\
  $\Rightarrow 4a = 24 \Rightarrow a = 6$\\
  and $(a - 3d)(a - d)(a + d)(a + 3d) = 945$\\
  $(a^2 - 9d^2)(a^2 - d^2) = 945$\\
  $(36 - 9d^2)(36 - d^2) = 945$\\
  $d^4 - 40d^2 + 144 = 105$\\
  $d^4 - 40d^2 + 39 = 0$\\
  $(d^2 - 1)(d^2 - 39) = 0$\\
  Since numbers are integers $\therefore d^2 \neq 39, \therefore d^2 = 1, d =
  \pm 1$\\
  Hence integers are, $3, 5, 7, 9$ or $9, 7, 4, 3$
\end{frame}
\begin{frame}{Problem 33}
  \textbf{33.} If the $p^{th}$ term of an A.P. is $q$ and the $q^{th}$ term is
  $p,$ find the first term and common difference. Also, show that $(p +
  q)^{th}$ term is zero.
\end{frame}
\begin{frame}{Solution of problem 33}
  \textbf{Solution:} Let $x$ be the first term and $d$ be the common
  difference.\\
  $t_p = x + (p - 1)d = q$\\
  $t_q = x + (q - 1)d = p$\\
  Subtracting these two we get, $q - p = (p - q)d \Rightarrow d = -1$\\
  Substituting this value for $t_p,$ we get $q = x + (p - 1). -1$\\
  $x = p + q - 1$\\
  $t_{p + q} = x + (p + q - 1)d = p + q -1 + (p + q - 1). -1 = 0$
\end{frame}
\begin{frame}{Problem 34}
  \textbf{34.} For an A.P. show that $t_m + t_{2n + m} = 2t_{m + n}.$
\end{frame}
\begin{frame}{Solution of problem 34}
  \textbf{Solution:} Let $a$ be the first term and $d$ be the common
  difference.\\
  $t_m = a + (m - 1)d$\\
  $t_{2n + m} = a + (2n + m - 1)d$\\
  $2t_{m+n} = 2[a + (m + n - 1)d]$\\
  $t_m + t_{2n + m} = 2a + (m - 1 + 2n + m - 1)d = 2[a + (m + n - 1)d] = 2t_{m
    + n}$
\end{frame}
\begin{frame}{Problem 35}
  \textbf{35.} Divide $15$ into three parts which are in A.P. and the sum of
  their squares is $83.$
\end{frame}
\begin{frame}{Solution of problem 35}
  \textbf{Solution:} Let $a - d, a , a + d$ be three such numbers.\\
  Given, $a - d + a + a + d = 15 \Rightarrow 3a = 15 \therefore a = 5$\\
  also, $(a - d)^2 + a^2 + (a + d)^2 = 83$\\
  $a^2 + d^2 - 2ad + a^2 + a^2 + d^2 + 2ad = 83$\\
  $3a^2 + 2d^2 = 83 \Rightarrow 3.5^2 + 2d^2 = 83$\\
  $2d^2 = 8 \Rightarrow d = \pm 2$\\
  Therefore, numbers are $3, 5, 7$ or $7, 5, 3$
\end{frame}
\begin{frame}{Problem 36}
  \textbf{36.} Three numbers are in A.P. Their sum is $27$ and the sum of their
  squares is $275.$ Find the numbers.
\end{frame}
\begin{frame}{Solution of problem 36}
  \textbf{Solution:} Let $a - d, a , a + d$ be three such numbers.\\
  Given, $a - d + a + a + d = 27 \Rightarrow 3a = 27 \therefore a = 9$\\
  also, $(a - d)^2 + a^2 + (a + d)^2 = 275$\\
  $a^2 + d^2 - 2ad + a^2 + a^2 + d^2 + 2ad = 275$\\
  $3a^2 + 2d^2 = 275 \Rightarrow 3.9^2 + 2d^2 = 275$\\
  $2d^2 = 32 \Rightarrow d = \pm 4$\\
  Therefore, numbers are $5, 9, 13$ or $13, 9, 5$
\end{frame}
\begin{frame}{Problem 37}
  \textbf{37.} The sum of three numbers in A.P. is $12$ and the sum of their
  cubes is $408,$ find them.
\end{frame}
\begin{frame}{Solution of problem 37}
  \textbf{Solution:} Let $a - d, a , a + d$ be three such numbers.\\
  Given, $a - d + a + a + d = 12 \Rightarrow 3a = 12 \therefore a = 4$\\
  also, $(a - d)^3 + a^3 + (a + d)^3 = 408$\\
  $a^3 - 3a^2d + 3ad^2 - d^3 + a^3 + a^3 + 3a^2d + 3ad^2 + d^3 = 408$\\
  $3a^3 + 6ad^2 = 408$\\
  $3.4^3 + 6.4.d^2 = 408$\\
  $24d^2 = 216 \Rightarrow d^2 = 9 \therefore d = \pm 3$\\
  Therefore, numbers are $1, 4, 7$ or $7, 4, 1$.
\end{frame}
\begin{frame}{Problem 38}
  \textbf{38.} Divide $20$ into four parts which are in A.P. such that the
  product of first and fourth is to product of second and third is $2:3.$
\end{frame}
\begin{frame}{Solution of problem 38}
  \textbf{Solution:} Let $a - 3d, a - d, a + d, a + 3d$ be the numbers.\\
  Given, $a - 3d + a - d + a + d + a + 3d = 20 \Rightarrow a = 5$\\
  also, $\frac{(a - 3d)(a + 3d)}{(a - d)(a + d)} = \frac{2}{3}$\\
  $\frac{a^2 - 9d^2}{a^2 - d^2} = \frac{2}{3}$\\
  $3a^2 - 27d^2 = 2a^2 - 2d^2$\\
  $a^2 = 25d^2 \Rightarrow d = \pm 1$\\
  Therfore, numbers are $2, 4, 6, 8$ or $8, 6, 4, 2$
\end{frame}
\begin{frame}{Problem 39}
  \textbf{39.} The sum of three numbers in A.P. is $-3$ and their product is
  $8.$ Find the numbers.
\end{frame}
\begin{frame}
  \textbf{Solution:} Let $a - d, a, a + d$ be the numbers.\\
  Given, $a - d + a + a + d = -3 \Rightarrow 3a = -3 \therefore a = -1$\\
  also, $(a - d)a(a + d) = 8$\\
  $a(a^2 - d^2) = 8$\\
  $-1(1 - d^2) = 8$\\
  $d^2 = 9 \Rightarrow d = \pm 3$\\
  Therefore, numbers are $-4, -1, 2$ or $2, -1, -4$
\end{frame}
\begin{frame}{Problem 40}
  \textbf{40.} Divide $32$ into four parts which are in A.P. such that the
  ratio of product of extremes to the product of means is $7:15.$
\end{frame}
\begin{frame}{Solution of problem 40}
  \textbf{Solution:} Let $a - 3d, a - d, a + d, a + 3d$ be the numbers.\\
  Given, $a - 3d + a - d + a + d + a + 3d = 32 \Rightarrow a = 8$\\
  also, $\frac{(a - 3d)(a + 3d)}{(a - d)(a + d)} = \frac{7}{15}$\\
  $\frac{a^2 - 9d^2}{a^2 - d^2} = \frac{7}{15}$\\
  $\frac{64 - 9d^2}{64 - d^2} = \frac{7}{15}$\\
  $960 - 135d^2 = 448 - 7d^2$\\
  $128d^2 = 512 \Rightarrow d^2 = 4 \therefore d = \pm 2$\\
  Therfore, numbers are $2, 6, 10, 14$ or $14, 10, 6, 2$
\end{frame}
\end{document}
